---
title: "Compare independent correlations: test coverage"
author: "Guillaume A. Rousselet"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: no
    number_sections: no
    toc: yes
    toc_depth: 2
---

Look at confidence interval coverage for Pearson's correlation using Zou's method and the percentile bootstrap. Do the same for Spearman's correlation using the bootstrap. 
Consider 3 cases:

- g = h = 0, difference = 0.1, vary rho
- g = 1, h = 0, difference = 0.1, vary rho
- rho = 0.3, difference = 0.2, vary g, h = 0

# Dependencies
```{r}
library(ggplot2)
library(tibble)
library(beepr)
source("./functions/theme_gar.txt")
## HC4 method to compare two independent Pearson's correlations
source("./functions/hc4.txt")
## g & h multivariate distributions
source("./functions/ghmult.txt")
## bootstrap + core corr functions
source("./functions/corfun.txt")
## meta-function for simulations
source("./functions/twocor.all.txt")
## plot functions
source("./functions/make_figures.txt")
```

# g = h = 0, difference = 0.1, vary rho
```{r}
rhoseq <- seq(0, 0.8, 0.1)
nseq <- seq(50, 500, 50)
nmax <- max(nseq)
nr <- length(rhoseq)
nsim <- 4000
es <- 0.1 # effect size
g <- 0
h <- 0
alpha <- .05
nboot <- 599

# coverage
cov.spear <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pears.zou <- array(0, dim = c(nsim, nr, length(nseq)))
# direction of discrepancy
dis.spear <- array(NA, dim = c(nsim, nr, length(nseq)))
dis.pears.boot <- array(NA, dim = c(nsim, nr, length(nseq)))
dis.pears.zou <- array(NA, dim = c(nsim, nr, length(nseq)))
# ci width
wid.spear <- array(0, dim = c(nsim, nr, length(nseq)))
wid.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
wid.pears.zou <- array(0, dim = c(nsim, nr, length(nseq)))

for(R in 1:nr){
  beep(2)
  set.seed(21) # set seed inside R loop to allow direct comparisons
  # (numbers are the same except for the correlation)
  rho <- rhoseq[R]
  print(paste("sim diff 0.1, g=0, h=0, rho =",rho))
  cmat1 <- matrix(c(1,rho,rho,1),2,2)
  cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)
  for(S in 1:nsim){
    if(S %% 500 == 0){print(paste0("sim ",S," / ",nsim,"..."))}
    # generate max n sample
    out1 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat1)  
    out2 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat2) 
    for(N in 1:length(nseq)){
      # subsample from max n sample
      x1 <- out1[1:nseq[N],1]
      y1 <- out1[1:nseq[N],2]
      x2 <- out2[1:nseq[N],1]
      y2 <- out2[1:nseq[N],2]

      # BOOTSTRAP SAMPLES -------------
      data1 <- matrix(sample(length(y1),size=length(y1)*nboot,replace=TRUE),nrow=nboot)
      data2 <- matrix(sample(length(y2),size=length(y2)*nboot,replace=TRUE),nrow=nboot)
      
      # PEARSON ---------------
      bvec1 <- apply(data1,1,pearsboot,x1,y1) # A 1 by nboot matrix.
      bvec2 <- apply(data2,1,pearsboot,x2,y2) # A 1 by nboot matrix.
      bvec <- bvec1 - bvec2
      ilow<-15
      ihi<-584
      if(length(y1)+length(y2) < 250){
        ilow<-14
        ihi<-585
      }
      if(length(y1)+length(y2) < 180){
        ilow<-11
        ihi<-588
      }
      if(length(y1)+length(y2) < 80){
        ilow<-8
        ihi<-592
      }
      if(length(y1)+length(y2) < 40){
        ilow<-7
        ihi<-593
      }
      bsort <- sort(bvec)
      ci <- c(bsort[ilow],bsort[ihi])
      if(-es >= ci[1] && -es <= ci[2]){
        cov.pears.boot[S, R, N] <- 1
      }
      if(cov.pears.boot[S, R, N] == 0){ # direction of discrepancy
        if(-es > ci[2]){
        dis.pears.boot[S, R, N] <- -1 # left shift
        }
        if(-es < ci[1]){
        dis.pears.boot[S, R, N] <- 1 # right shift
        }
      }
      wid.pears.boot[S, R, N] <- ci[2] - ci[1] # ci width
      
      # ZOU 2007 --------------
      r1 <- cor(x1, y1)
      r2 <- cor(x2, y2)
      n1 <- length(x1)
      n2 <- length(x2)
      ci <- r.ind.ci(r1, r2, n1, n2)
      if(-es >= ci[1] && -es <= ci[2]){
        cov.pears.zou[S, R, N] <- 1
      }
      if(cov.pears.zou[S, R, N] == 0){ # direction of discrepancy
        if(-es > ci[2]){
        dis.pears.zou[S, R, N] <- -1 # left shift
        }
        if(-es < ci[1]){
        dis.pears.zou[S, R, N] <- 1 # right shift
        }
      }
      wid.pears.zou[S, R, N] <- ci[2] - ci[1] # ci width
      
      # SPEARMAN --------------
      bvec1 <- apply(data1,1,spearboot,x1,y1) # A 1 by nboot matrix.
      bvec2 <- apply(data2,1,spearboot,x2,y2) # A 1 by nboot matrix.
      bvec <- bvec1 - bvec2
      bsort <- sort(bvec)
      ilow <- round((alpha/2) * nboot)
      ihi <- nboot - ilow
      ilow <- ilow+1
      ci <- c(bsort[ilow],bsort[ihi])
      if(-es >= ci[1] && -es <= ci[2]){
        cov.spear[S, R, N] <- 1
      }
      if(cov.spear[S, R, N] == 0){ # direction of discrepancy
        if(-es > ci[2]){
        dis.spear[S, R, N] <- -1 # left shift
        }
        if(-es < ci[1]){
        dis.spear[S, R, N] <- 1 # right shift
        }
      }
      wid.spear[S, R, N] <- ci[2] - ci[1] # ci width
    }
  }
}

beep(8)

save(cov.spear, cov.pears.boot, cov.pears.zou,
     dis.spear, dis.pears.boot, dis.pears.zou,
     wid.spear, wid.pears.boot, wid.pears.zou,
     rhoseq, nseq, nsim, es, g, h,
     file = "./data/diff01_g0h0_varyrho_covtest.RData")
```

## Pearson + Zou
```{r}
load("./data/diff01_g0h0_varyrho_covtest.RData")
pres <- apply(cov.pears, c(2,3), mean)
df <- mdf(pres, nseq, rhoseq)
p <- plot.res.tp(df, ylim = c(0, 1))
p.pears <- p + ggtitle("Pearson + Zou") +
  theme(legend.position = "right") +
  labs(y = "Coverage prob.") + 
  geom_hline(yintercept = 0.95) +
  coord_cartesian(ylim = c(0.5, 1))
p.pears
```

## Pearson + bootstrap
```{r}
pres <- apply(cov.pears, c(2,3), mean)
df <- mdf(pres, nseq, rhoseq)
p <- plot.res.tp(df, ylim = c(0, 1))
p.pears <- p + ggtitle("Pearson + bootstrap") +
  theme(legend.position = "right") +
  labs(y = "Coverage prob.") + 
  geom_hline(yintercept = 0.95) +
  coord_cartesian(ylim = c(0.5, 1))
p.pears
```

## Spearman + bootstrap
```{r}
pres <- apply(cov.spear, c(2,3), mean)
df <- mdf(pres, nseq, rhoseq)
p <- plot.res.tp(df, ylim = c(0, 1))
p.spear <- p + ggtitle("Spearman + bootstrap") +
  theme(legend.position = "none") +
  labs(y = "Coverage prob.") + 
  geom_hline(yintercept = 0.95) +
  coord_cartesian(ylim = c(0.5, 1))
p.spear
```

# g = 1, h = 0, vary rho
```{r}
rhoseq <- seq(0, 0.8, 0.1)
nseq <- seq(50, 500, 50)
nmax <- max(nseq)
nr <- length(rhoseq)
nsim <- 4000
es <- 0.1 # effect size
g <- 1
h <- 0
alpha <- .05
nboot <- 599

# coverage
cov.spear <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pears.zou <- array(0, dim = c(nsim, nr, length(nseq)))
# direction of discrepancy
dis.spear <- array(NA, dim = c(nsim, nr, length(nseq)))
dis.pears.boot <- array(NA, dim = c(nsim, nr, length(nseq)))
dis.pears.zou <- array(NA, dim = c(nsim, nr, length(nseq)))
# ci width
wid.spear <- array(0, dim = c(nsim, nr, length(nseq)))
wid.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
wid.pears.zou <- array(0, dim = c(nsim, nr, length(nseq)))

for(R in 1:nr){
  beep(2)
  set.seed(21) # set seed inside R loop to allow direct comparisons
  # (numbers are the same except for the correlation)
  rho <- rhoseq[R]
  print(paste("sim diff 0.1, g=0, h=0, rho =",rho))
  cmat1 <- matrix(c(1,rho,rho,1),2,2)
  cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)
  for(S in 1:nsim){
    if(S %% 500 == 0){print(paste0("sim ",S," / ",nsim,"..."))}
    # generate max n sample
    out1 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat1)  
    out2 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat2) 
    for(N in 1:length(nseq)){
      # subsample from max n sample
      x1 <- out1[1:nseq[N],1]
      y1 <- out1[1:nseq[N],2]
      x2 <- out2[1:nseq[N],1]
      y2 <- out2[1:nseq[N],2]

      # BOOTSTRAP SAMPLES -------------
      data1 <- matrix(sample(length(y1),size=length(y1)*nboot,replace=TRUE),nrow=nboot)
      data2 <- matrix(sample(length(y2),size=length(y2)*nboot,replace=TRUE),nrow=nboot)
      
      # PEARSON ---------------
      bvec1 <- apply(data1,1,pearsboot,x1,y1) # A 1 by nboot matrix.
      bvec2 <- apply(data2,1,pearsboot,x2,y2) # A 1 by nboot matrix.
      bvec <- bvec1 - bvec2
      ilow<-15
      ihi<-584
      if(length(y1)+length(y2) < 250){
        ilow<-14
        ihi<-585
      }
      if(length(y1)+length(y2) < 180){
        ilow<-11
        ihi<-588
      }
      if(length(y1)+length(y2) < 80){
        ilow<-8
        ihi<-592
      }
      if(length(y1)+length(y2) < 40){
        ilow<-7
        ihi<-593
      }
      bsort <- sort(bvec)
      ci <- c(bsort[ilow],bsort[ihi])
      if(-es >= ci[1] && -es <= ci[2]){
        cov.pears.boot[S, R, N] <- 1
      }
      if(cov.pears.boot[S, R, N] == 0){ # direction of discrepancy
        if(-es > ci[2]){
        dis.pears.boot[S, R, N] <- -1 # left shift
        }
        if(-es < ci[1]){
        dis.pears.boot[S, R, N] <- 1 # right shift
        }
      }
      wid.pears.boot[S, R, N] <- ci[2] - ci[1] # ci width
      
      # ZOU 2007 --------------
      r1 <- cor(x1, y1)
      r2 <- cor(x2, y2)
      n1 <- length(x1)
      n2 <- length(x2)
      ci <- r.ind.ci(r1, r2, n1, n2)
      if(-es >= ci[1] && -es <= ci[2]){
        cov.pears.zou[S, R, N] <- 1
      }
      if(cov.pears.zou[S, R, N] == 0){ # direction of discrepancy
        if(-es > ci[2]){
        dis.pears.zou[S, R, N] <- -1 # left shift
        }
        if(-es < ci[1]){
        dis.pears.zou[S, R, N] <- 1 # right shift
        }
      }
      wid.pears.zou[S, R, N] <- ci[2] - ci[1] # ci width
      
      # SPEARMAN --------------
      bvec1 <- apply(data1,1,spearboot,x1,y1) # A 1 by nboot matrix.
      bvec2 <- apply(data2,1,spearboot,x2,y2) # A 1 by nboot matrix.
      bvec <- bvec1 - bvec2
      bsort <- sort(bvec)
      ilow <- round((alpha/2) * nboot)
      ihi <- nboot - ilow
      ilow <- ilow+1
      ci <- c(bsort[ilow],bsort[ihi])
      if(-es >= ci[1] && -es <= ci[2]){
        cov.spear[S, R, N] <- 1
      }
      if(cov.spear[S, R, N] == 0){ # direction of discrepancy
        if(-es > ci[2]){
        dis.spear[S, R, N] <- -1 # left shift
        }
        if(-es < ci[1]){
        dis.spear[S, R, N] <- 1 # right shift
        }
      }
      wid.spear[S, R, N] <- ci[2] - ci[1] # ci width
    }
  }
}

beep(8)

save(cov.spear, cov.pears.boot, cov.pears.zou,
     dis.spear, dis.pears.boot, dis.pears.zou,
     wid.spear, wid.pears.boot, wid.pears.zou,
     rhoseq, nseq, nsim, es, g, h,
     file = "./data/diff01_g1h0_varyrho_covtest.RData")
```

## Results: coverage

### Pearson + Zou
```{r}
load("./data/diff01_g1h0_varyrho_covtest.RData")
pres <- apply(cov.pears.zou, c(2,3), mean)
df <- mdf(pres, nseq, rhoseq)
p <- plot.res.tp(df, ylim = c(0, 1))
p <- p + ggtitle("Pearson + Zou") +
  theme(legend.position = "right") +
  labs(y = "Coverage prob.") + 
  geom_hline(yintercept = 0.95) +
  coord_cartesian(ylim = c(0.45, 1))
p
```

### Pearson + bootstrap
```{r}
pres <- apply(cov.pears.boot, c(2,3), mean)
df <- mdf(pres, nseq, rhoseq)
p <- plot.res.tp(df, ylim = c(0, 1))
p <- p + ggtitle("Pearson + bootstrap") +
  theme(legend.position = "right") +
  labs(y = "Coverage prob.") + 
  geom_hline(yintercept = 0.95) +
  coord_cartesian(ylim = c(0.45, 1))
p
```

### Spearman + bootstrap
```{r}
pres <- apply(cov.spear, c(2,3), mean)
df <- mdf(pres, nseq, rhoseq)
p <- plot.res.tp(df, ylim = c(0, 1))
p <- p + ggtitle("Spearman + bootstrap") +
  theme(legend.position = "none") +
  labs(y = "Coverage prob.") + 
  geom_hline(yintercept = 0.95) +
  coord_cartesian(ylim = c(0.45, 1))
p
```

## Results: width

### Pearson + Zou
```{r}
pres <- apply(wid.pears.zou, c(2,3), median)
df <- mdf(pres, nseq, rhoseq)
p <- plot.res.tp(df, ylim = c(0, 1))
p + ggtitle("Pearson + Zou") +
  theme(legend.position = "right") +
  labs(y = "CI width") #+ 
  # geom_hline(yintercept = 0.95) +
  # coord_cartesian(ylim = c(0.45, 1))
p
```

### Pearson + bootstrap
```{r}
pres <- apply(wid.pears.boot, c(2,3), median)
df <- mdf(pres, nseq, rhoseq)
p <- plot.res.tp(df, ylim = c(0, 1))
p <- p + ggtitle("Pearson + bootstrap") +
  theme(legend.position = "right") +
  labs(y = "CI width") #+ 
  # geom_hline(yintercept = 0.95) +
  # coord_cartesian(ylim = c(0.45, 1))
p
```

### Spearman + bootstrap
```{r}
pres <- apply(wid.spear.boot, c(2,3), median)
df <- mdf(pres, nseq, rhoseq)
p <- plot.res.tp(df, ylim = c(0, 1))
p <- p + ggtitle("Spearman + bootstrap") +
  theme(legend.position = "none") +
  labs(y = "CI width") #+ 
  # geom_hline(yintercept = 0.95) +
  # coord_cartesian(ylim = c(0.45, 1))
p
```

## Results: direction of discrepancy

### Pearson + Zou
```{r}
pres <- apply(dis.pears.zou, c(2,3), mean, na.rm = TRUE)
df <- mdf(pres, nseq, rhoseq)
p <- plot.res.tp(df, ylim = c(0, 1))
p + ggtitle("Pearson + Zou") +
  theme(legend.position = "right") +
  labs(y = "CI width") + 
  geom_hline(yintercept = 0) #+
  # coord_cartesian(ylim = c(0.45, 1))
p
```

### Pearson + bootstrap
```{r}
pres <- apply(dis.pears.boot, c(2,3), mean, na.rm = TRUE)
df <- mdf(pres, nseq, rhoseq)
p <- plot.res.tp(df, ylim = c(0, 1))
p <- p + ggtitle("Pearson + bootstrap") +
  theme(legend.position = "right") +
  labs(y = "CI width") + 
  geom_hline(yintercept = 0) #+
  # coord_cartesian(ylim = c(0.45, 1))
p
```

### Spearman + bootstrap
```{r}
pres <- apply(dis.spear.boot, c(2,3), mean, na.rm = TRUE)
df <- mdf(pres, nseq, rhoseq)
p <- plot.res.tp(df, ylim = c(0, 1))
p <- p + ggtitle("Spearman + bootstrap") +
  theme(legend.position = "none") +
  labs(y = "CI width") + 
  geom_hline(yintercept = 0) #+
  # coord_cartesian(ylim = c(0.45, 1))
p
```

# rho = 0.3, difference = 0.2, vary g
```{r}
gseq <- seq(0, 1, 0.1)
ng <- length(gseq)
nseq <- seq(50, 500, 50)
nmax <- max(nseq)
nsim <- 4000
rho = 0.3
es <- 0.2 # effect size
h <- 0
cmat1 <- matrix(c(1,rho,rho,1),2,2)
cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)
alpha <- .05
nboot <- 599

# coverage
cov.spear <- array(0, dim = c(nsim, ng, length(nseq)))
cov.pears.boot <- array(0, dim = c(nsim, ng, length(nseq)))
cov.pears.zou <- array(0, dim = c(nsim, ng, length(nseq)))
# direction of discrepancy
dis.spear <- array(NA, dim = c(nsim, ng, length(nseq)))
dis.pears.boot <- array(NA, dim = c(nsim, ng, length(nseq)))
dis.pears.zou <- array(NA, dim = c(nsim, ng, length(nseq)))
# ci width
wid.spear <- array(0, dim = c(nsim, ng, length(nseq)))
wid.pears.boot <- array(0, dim = c(nsim, ng, length(nseq)))
wid.pears.zou <- array(0, dim = c(nsim, ng, length(nseq)))

for(G in 1:ng){
  beep(2)
  set.seed(21) # set seed inside R loop to allow direct comparisons
  # (numbers are the same except for the correlation)
  g <- gseq[G]
  print(paste("sim diff 0.1, g=0, h=0, rho =",rho))
  for(S in 1:nsim){
    if(S %% 500 == 0){print(paste0("sim ",S," / ",nsim,"..."))}
    # generate max n sample
    out1 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat1)  
    out2 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat2) 
    for(N in 1:length(nseq)){
      # subsample from max n sample
      x1 <- out1[1:nseq[N],1]
      y1 <- out1[1:nseq[N],2]
      x2 <- out2[1:nseq[N],1]
      y2 <- out2[1:nseq[N],2]

      # BOOTSTRAP SAMPLES -------------
      data1 <- matrix(sample(length(y1),size=length(y1)*nboot,replace=TRUE),nrow=nboot)
      data2 <- matrix(sample(length(y2),size=length(y2)*nboot,replace=TRUE),nrow=nboot)
      
      # PEARSON ---------------
      bvec1 <- apply(data1,1,pearsboot,x1,y1) # A 1 by nboot matrix.
      bvec2 <- apply(data2,1,pearsboot,x2,y2) # A 1 by nboot matrix.
      bvec <- bvec1 - bvec2
      ilow<-15
      ihi<-584
      if(length(y1)+length(y2) < 250){
        ilow<-14
        ihi<-585
      }
      if(length(y1)+length(y2) < 180){
        ilow<-11
        ihi<-588
      }
      if(length(y1)+length(y2) < 80){
        ilow<-8
        ihi<-592
      }
      if(length(y1)+length(y2) < 40){
        ilow<-7
        ihi<-593
      }
      bsort <- sort(bvec)
      ci <- c(bsort[ilow],bsort[ihi])
      if(-es >= ci[1] && -es <= ci[2]){
        cov.pears.boot[S, G, N] <- 1
      }
      if(cov.pears.boot[S, G, N] == 0){ # direction of discrepancy
        if(-es > ci[2]){
        dis.pears.boot[S, G, N] <- -1 # left shift
        }
        if(-es < ci[1]){
        dis.pears.boot[S, G, N] <- 1 # right shift
        }
      }
      wid.pears.boot[S, G, N] <- ci[2] - ci[1] # ci width
      
      # ZOU 2007 --------------
      r1 <- cor(x1, y1)
      r2 <- cor(x2, y2)
      n1 <- length(x1)
      n2 <- length(x2)
      ci <- r.ind.ci(r1, r2, n1, n2)
      if(-es >= ci[1] && -es <= ci[2]){
        cov.pears.zou[S, G, N] <- 1
      }
      if(cov.pears.zou[S, G, N] == 0){ # direction of discrepancy
        if(-es > ci[2]){
        dis.pears.zou[S, G, N] <- -1 # left shift
        }
        if(-es < ci[1]){
        dis.pears.zou[S, G, N] <- 1 # right shift
        }
      }
      wid.pears.zou[S, G, N] <- ci[2] - ci[1] # ci width
      
      # SPEARMAN --------------
      bvec1 <- apply(data1,1,spearboot,x1,y1) # A 1 by nboot matrix.
      bvec2 <- apply(data2,1,spearboot,x2,y2) # A 1 by nboot matrix.
      bvec <- bvec1 - bvec2
      bsort <- sort(bvec)
      ilow <- round((alpha/2) * nboot)
      ihi <- nboot - ilow
      ilow <- ilow+1
      ci <- c(bsort[ilow],bsort[ihi])
      if(-es >= ci[1] && -es <= ci[2]){
        cov.spear[S, G, N] <- 1
      }
      if(cov.spear[S, G, N] == 0){ # direction of discrepancy
        if(-es > ci[2]){
        dis.spear[S, G, N] <- -1 # left shift
        }
        if(-es < ci[1]){
        dis.spear[S, G, N] <- 1 # right shift
        }
      }
      wid.spear[S, G, N] <- ci[2] - ci[1] # ci width
    }
  }
}

beep(8)

save(cov.spear, cov.pears.boot, cov.pears.zou,
     dis.spear, dis.pears.boot, dis.pears.zou,
     wid.spear, wid.pears.boot, wid.pears.zou,
     rhoseq, nseq, nsim, es, g, h,
     file = "./data/rho03_diff02_h0_vg_covtest.RData")
```

