---
title: "Compare independent correlations: 0.1 difference, vary rho"
author: "Guillaume A. Rousselet"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: no
    number_sections: no
    toc: yes
    toc_depth: 2
---

# Dependencies
```{r}
library(ggplot2)
library(tibble)
library(beepr)
source("./functions/theme_gar.txt")
## HC4 method to compare two independent Pearson's correlations
source("./functions/hc4.txt")
## g & h multivariate distributions
source("./functions/ghmult.txt")
## bootstrap + core corr functions
source("./functions/corfun.txt")
## meta-function for simulations
source("./functions/twocor.all.txt")
## plot functions
source("./functions/make_figures.txt")
```

```{r}
sessionInfo()
```

# g=0, h=0
```{r, eval = FALSE}
rhoseq <- seq(0, 0.8, 0.1)
nseq <- seq(50, 500, 50)
nmax <- max(nseq)
nr <- length(rhoseq)
nsim <- 5000
es <- 0.1 # effect size
g <- 0
h <- 0
res.wcorr <- array(0, dim = c(nsim, nr, length(nseq)))
res.pbend <- array(0, dim = c(nsim, nr, length(nseq)))
res.spear <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, nr, length(nseq)))

cov.wcorr <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pbend <- array(0, dim = c(nsim, nr, length(nseq)))
cov.spear <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pears <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pears.zou <- array(0, dim = c(nsim, nr, length(nseq)))

for(R in 1:nr){
  beep(2)
  set.seed(21) # set seed inside R loop to allow direct comparisons
  # (numbers are the same except for the correlation)
  rho <- rhoseq[R]
  print(paste("sim diff 0.1, g=0, h=0, rho =",rho))
  cmat1 <- matrix(c(1,rho,rho,1),2,2)
  cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)
  for(S in 1:nsim){
    if(S %% 1000 == 0){print(paste0("sim ",S," / ",nsim,"..."))}
    # generate max n sample
    out1 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat1)  
    out2 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat2) 
    for(N in 1:length(nseq)){
      # subsample from max n sample
      res <- twocor.all(out1[1:nseq[N],1],out1[1:nseq[N],2],out2[1:nseq[N],1],out2[1:nseq[N],2],
                        rho = -es) # rho1 minus rho2 = -es
      res.wcorr[S, R, N] <- res$sig.boot.wcorr
      res.pbend[S, R, N] <- res$sig.boot.pbcor
      res.spear[S, R, N] <- res$sig.boot.spear
      res.pears.boot[S, R, N] <- res$sig.boot.pears
      res.pears.fish[S, R, N] <- res$sig.fisher.pears
      res.pears.hc4[S, R, N] <- res$sig.hc4.pears
      
      cov.wcorr[S, R, N] <- res$cov.boot.wcorr  
      cov.pbend[S, R, N] <- res$cov.boot.pbcor
      cov.spear[S, R, N] <- res$cov.boot.spear 
      cov.pears[S, R, N] <- res$cov.boot.pears
      cov.pears.zou[S, R, N] <- res$cov.zou.pears
    }
  }
save(res.spear, res.pears.boot, res.pears.fish, 
     res.pears.hc4, res.wcorr, res.pbend,
     cov.wcorr, cov.pbend, cov.spear, cov.pears, cov.pears.zou
     rhoseq, nseq, nsim, es, g, h,
     file = "./data/diff01_g0h0_varyrho.RData")
}
beep(8)
```

## Results: true positives

### Load data + generate all plot objects
```{r}
load("./data/diff01_g0h0_varyrho.RData")
PO <- plot.all.tp(ylim = c(0, 1))
```

### Pearson + Fisher
```{r}
pA <- PO$p.pears.fish # + guides(col = guide_legend(ncol = 3, override.aes = list(size = 3)))
pA
```

### Pearson + bootstrap
```{r}
pB <- PO$p.pears.boot
pB
```

### Pearson + HC4
```{r}
pC <- PO$p.pears.hc4
pC
```

### Spearman + bootstrap
```{r}
pD <- PO$p.spear
pD
```

### Percentage bend + bootstrap
```{r}
pE <- PO$p.pbend
pE
```

### Winsorized correlation + bootstrap
```{r}
pF <- PO$p.wcorr
pF
```

### Combine panels

```{r, eval = FALSE}
legend <- cowplot::get_legend(pA)

fig <- cowplot::plot_grid(pA + theme(legend.position = "none"), pD,
                          pB, pE,  
                          pC, pF,
                          labels = c("A", "B", "C", "D", "E", "F"),
                          nrow = 3)

cowplot::plot_grid(fig, legend,
                   ncol = 2,
                   rel_widths = c(3, .3))

ggsave("./figures/diff01_g0h0_vrho_power.pdf", width = 30, height = 30, units = "cm")
```

## Results: coverage

### Load data + generate all plot objects
```{r}
# load("./data/diff01_g0h0_varyrho.RData")
PO <- plot.all.cov(ylim = c(0, 1))
```

### Pearson + bootstrap
```{r}
pA <- PO$p.pears # + guides(col = guide_legend(ncol = 3, override.aes = list(size = 3)))
pA
```

### Spearman + bootstrap
```{r}
pB <- PO$p.spear
pB
```

### Percentage bend + bootstrap
```{r}
pC <- PO$p.pbend
pC
```

### Winsorized correlation + bootstrap
```{r}
pD <- PO$p.wcorr
pD
```

### Combine panels

```{r, eval = FALSE}
legend <- cowplot::get_legend(pA)

fig <- cowplot::plot_grid(pA + theme(legend.position = "none"), pB, pC, pD,
                          labels = c("A", "B", "C", "D"),
                          nrow = 2)

cowplot::plot_grid(fig, legend,
                   ncol = 2,
                   rel_widths = c(3, .4))

ggsave("./figures/diff01_g0h0_vrho_coverage.pdf", width = 30, height = 20, units = "cm")
```

# g=0, h=0.2
```{r, eval = FALSE}
rhoseq <- seq(0, 0.8, 0.1)
nseq <- seq(50, 500, 50)
nmax <- max(nseq)
nr <- length(rhoseq)
nsim <- 5000
es <- 0.1 # effect size
g <- 0
h <- 0.2
res.wcorr <- array(0, dim = c(nsim, nr, length(nseq)))
res.pbend <- array(0, dim = c(nsim, nr, length(nseq)))
res.spear <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, nr, length(nseq)))

cov.wcorr <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pbend <- array(0, dim = c(nsim, nr, length(nseq)))
cov.spear <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pears <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pears.zou <- array(0, dim = c(nsim, nr, length(nseq)))

for(R in 1:nr){
  beep(2)
  set.seed(21) # set seed inside R loop to allow direct comparisons
  # (numbers are the same except for the correlation)
  rho <- rhoseq[R]
  print(paste0("sim diff=0.1, g=0, h=0.2, rho=",rho))
  cmat1 <- matrix(c(1,rho,rho,1),2,2)
  cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)
  for(S in 1:nsim){
    if(S %% 1000 == 0){print(paste0("sim ",S," / ",nsim,"..."))}
    # generate max n sample
    out1 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat1)  
    out2 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat2) 
    for(N in 1:length(nseq)){
      # subsample from max n sample
      res <- twocor.all(out1[1:nseq[N],1],out1[1:nseq[N],2],out2[1:nseq[N],1],out2[1:nseq[N],2],
                        rho = -es) # rho1 minus rho2 = -es
      res.wcorr[S, R, N] <- res$sig.boot.wcorr
      res.pbend[S, R, N] <- res$sig.boot.pbcor
      res.spear[S, R, N] <- res$sig.boot.spear
      res.pears.boot[S, R, N] <- res$sig.boot.pears
      res.pears.fish[S, R, N] <- res$sig.fisher.pears
      res.pears.hc4[S, R, N] <- res$sig.hc4.pears
      
      cov.wcorr[S, R, N] <- res$cov.boot.wcorr  
      cov.pbend[S, R, N] <- res$cov.boot.pbcor
      cov.spear[S, R, N] <- res$cov.boot.spear 
      cov.pears[S, R, N] <- res$cov.boot.pears
      cov.pears.zou[S, R, N] <- res$cov.zou.pears
    }
  }
save(res.spear, res.pears.boot, res.pears.fish, 
     res.pears.hc4, res.wcorr, res.pbend,
     cov.wcorr, cov.pbend, cov.spear, cov.pears, cov.pears.zou
     rhoseq, nseq, nsim, es, g, h,
     file = "./data/diff01_g0h02_varyrho.RData")
}
beep(8)
```

# g=0.2, h=0
```{r, eval = FALSE}
rhoseq <- seq(0, 0.8, 0.1)
nseq <- seq(50, 500, 50)
nmax <- max(nseq)
nr <- length(rhoseq)
nsim <- 5000
es <- 0.1 # effect size
g <- 0.2
h <- 0
res.wcorr <- array(0, dim = c(nsim, nr, length(nseq)))
res.pbend <- array(0, dim = c(nsim, nr, length(nseq)))
res.spear <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, nr, length(nseq)))

cov.wcorr <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pbend <- array(0, dim = c(nsim, nr, length(nseq)))
cov.spear <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pears <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pears.zou <- array(0, dim = c(nsim, nr, length(nseq)))

for(R in 1:nr){
  beep(2)
  set.seed(21) # set seed inside R loop to allow direct comparisons
  # (numbers are the same except for the correlation)
  rho <- rhoseq[R]
  print(paste0("sim diff=0.1, g=0.2, h=0, rho=",rho))
  cmat1 <- matrix(c(1,rho,rho,1),2,2)
  cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)
  for(S in 1:nsim){
    if(S %% 1000 == 0){print(paste0("sim ",S," / ",nsim,"..."))}
    # generate max n sample
    out1 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat1)  
    out2 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat2) 
    for(N in 1:length(nseq)){
      # subsample from max n sample
      res <- twocor.all(out1[1:nseq[N],1],out1[1:nseq[N],2],out2[1:nseq[N],1],out2[1:nseq[N],2],
                        rho = -es) # rho1 minus rho2 = -es
      res.wcorr[S, R, N] <- res$sig.boot.wcorr
      res.pbend[S, R, N] <- res$sig.boot.pbcor
      res.spear[S, R, N] <- res$sig.boot.spear
      res.pears.boot[S, R, N] <- res$sig.boot.pears
      res.pears.fish[S, R, N] <- res$sig.fisher.pears
      res.pears.hc4[S, R, N] <- res$sig.hc4.pears
      
      cov.wcorr[S, R, N] <- res$cov.boot.wcorr  
      cov.pbend[S, R, N] <- res$cov.boot.pbcor
      cov.spear[S, R, N] <- res$cov.boot.spear 
      cov.pears[S, R, N] <- res$cov.boot.pears
      cov.pears.zou[S, R, N] <- res$cov.zou.pears
    }
  }
save(res.spear, res.pears.boot, res.pears.fish, 
     res.pears.hc4, res.wcorr, res.pbend,
     cov.wcorr, cov.pbend, cov.spear, cov.pears, cov.pears.zou
     rhoseq, nseq, nsim, es, g, h,
     file = "./data/diff01_g02h0_varyrho.RData")
}
beep(8)
```

# g=0.2, h=0.2
```{r, eval = FALSE}
rhoseq <- seq(0, 0.8, 0.1)
nseq <- seq(50, 500, 50)
nmax <- max(nseq)
nr <- length(rhoseq)
nsim <- 5000
es <- 0.1 # effect size
g <- 0.2
h <- 0.2
res.wcorr <- array(0, dim = c(nsim, nr, length(nseq)))
res.pbend <- array(0, dim = c(nsim, nr, length(nseq)))
res.spear <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, nr, length(nseq)))

cov.wcorr <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pbend <- array(0, dim = c(nsim, nr, length(nseq)))
cov.spear <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pears <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pears.zou <- array(0, dim = c(nsim, nr, length(nseq)))

for(R in 1:nr){
  beep(2)
  set.seed(21) # set seed inside R loop to allow direct comparisons
  # (numbers are the same except for the correlation)
  rho <- rhoseq[R]
  print(paste0("sim diff=0.1, g=0.2, h=0.2, rho=",rho))
  cmat1 <- matrix(c(1,rho,rho,1),2,2)
  cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)
  for(S in 1:nsim){
    if(S %% 1000 == 0){print(paste0("sim ",S," / ",nsim,"..."))}
    # generate max n sample
    out1 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat1)  
    out2 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat2) 
    for(N in 1:length(nseq)){
      # subsample from max n sample
      res <- twocor.all(out1[1:nseq[N],1],out1[1:nseq[N],2],out2[1:nseq[N],1],out2[1:nseq[N],2],
                        rho = -es) # rho1 minus rho2 = -es
      res.wcorr[S, R, N] <- res$sig.boot.wcorr
      res.pbend[S, R, N] <- res$sig.boot.pbcor
      res.spear[S, R, N] <- res$sig.boot.spear
      res.pears.boot[S, R, N] <- res$sig.boot.pears
      res.pears.fish[S, R, N] <- res$sig.fisher.pears
      res.pears.hc4[S, R, N] <- res$sig.hc4.pears
      
      cov.wcorr[S, R, N] <- res$cov.boot.wcorr  
      cov.pbend[S, R, N] <- res$cov.boot.pbcor
      cov.spear[S, R, N] <- res$cov.boot.spear 
      cov.pears[S, R, N] <- res$cov.boot.pears
      cov.pears.zou[S, R, N] <- res$cov.zou.pears
    }
  }
save(res.spear, res.pears.boot, res.pears.fish, 
     res.pears.hc4, res.wcorr, res.pbend,
     cov.wcorr, cov.pbend, cov.spear, cov.pears, cov.pears.zou
     rhoseq, nseq, nsim, es, g, h,
     file = "./data/diff01_g02h02_varyrho.RData")
}
beep(8)
```

# g=1, h=0
```{r, eval = FALSE}
rhoseq <- seq(0, 0.8, 0.1)
nseq <- seq(50, 500, 50)
nmax <- max(nseq)
nr <- length(rhoseq)
nsim <- 5000
es <- 0.1 # effect size
g <- 1
h <- 0
res.wcorr <- array(0, dim = c(nsim, nr, length(nseq)))
res.pbend <- array(0, dim = c(nsim, nr, length(nseq)))
res.spear <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, nr, length(nseq)))

cov.wcorr <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pbend <- array(0, dim = c(nsim, nr, length(nseq)))
cov.spear <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pears <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pears.zou <- array(0, dim = c(nsim, nr, length(nseq)))

for(R in 1:nr){
  beep(2)
  set.seed(21) # set seed inside R loop to allow direct comparisons
  # (numbers are the same except for the correlation)
  rho <- rhoseq[R]
  print(paste0("sim diff=0.1, g=1, h=0, rho=",rho))
  cmat1 <- matrix(c(1,rho,rho,1),2,2)
  cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)
  for(S in 1:nsim){
    if(S %% 1000 == 0){print(paste0("sim ",S," / ",nsim,"..."))}
    # generate max n sample
    out1 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat1)  
    out2 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat2) 
    for(N in 1:length(nseq)){
      # subsample from max n sample
      res <- twocor.all(out1[1:nseq[N],1],out1[1:nseq[N],2],out2[1:nseq[N],1],out2[1:nseq[N],2],
                        rho = -es) # rho1 minus rho2 = -es
      res.wcorr[S, R, N] <- res$sig.boot.wcorr
      res.pbend[S, R, N] <- res$sig.boot.pbcor
      res.spear[S, R, N] <- res$sig.boot.spear
      res.pears.boot[S, R, N] <- res$sig.boot.pears
      res.pears.fish[S, R, N] <- res$sig.fisher.pears
      res.pears.hc4[S, R, N] <- res$sig.hc4.pears
      
      cov.wcorr[S, R, N] <- res$cov.boot.wcorr  
      cov.pbend[S, R, N] <- res$cov.boot.pbcor
      cov.spear[S, R, N] <- res$cov.boot.spear 
      cov.pears[S, R, N] <- res$cov.boot.pears
      cov.pears.zou[S, R, N] <- res$cov.zou.pears
    }
  }
save(res.spear, res.pears.boot, res.pears.fish, 
     res.pears.hc4, res.wcorr, res.pbend,
     cov.wcorr, cov.pbend, cov.spear, cov.pears, cov.pears.zou
     rhoseq, nseq, nsim, es, g, h,
     file = "./data/diff01_g1h0_varyrho.RData")
}
beep(8)
```

# g=1, h=0.2
```{r, eval = FALSE}
rhoseq <- seq(0, 0.8, 0.1)
nseq <- seq(50, 500, 50)
nmax <- max(nseq)
nr <- length(rhoseq)
nsim <- 5000
es <- 0.1 # effect size
g <- 1
h <- 0.2
res.wcorr <- array(0, dim = c(nsim, nr, length(nseq)))
res.pbend <- array(0, dim = c(nsim, nr, length(nseq)))
res.spear <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, nr, length(nseq)))

cov.wcorr <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pbend <- array(0, dim = c(nsim, nr, length(nseq)))
cov.spear <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pears <- array(0, dim = c(nsim, nr, length(nseq)))
cov.pears.zou <- array(0, dim = c(nsim, nr, length(nseq)))

for(R in 1:nr){
  beep(2)
  set.seed(21) # set seed inside R loop to allow direct comparisons
  # (numbers are the same except for the correlation)
  rho <- rhoseq[R]
  print(paste0("sim diff=0.1, g=1, h=0.2, rho=",rho))
  cmat1 <- matrix(c(1,rho,rho,1),2,2)
  cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)
  for(S in 1:nsim){
    if(S %% 1000 == 0){print(paste0("sim ",S," / ",nsim,"..."))}
    # generate max n sample
    out1 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat1)  
    out2 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat2) 
    for(N in 1:length(nseq)){
      # subsample from max n sample
      res <- twocor.all(out1[1:nseq[N],1],out1[1:nseq[N],2],out2[1:nseq[N],1],out2[1:nseq[N],2],
                        rho = -es) # rho1 minus rho2 = -es
      res.wcorr[S, R, N] <- res$sig.boot.wcorr
      res.pbend[S, R, N] <- res$sig.boot.pbcor
      res.spear[S, R, N] <- res$sig.boot.spear
      res.pears.boot[S, R, N] <- res$sig.boot.pears
      res.pears.fish[S, R, N] <- res$sig.fisher.pears
      res.pears.hc4[S, R, N] <- res$sig.hc4.pears
      
      cov.wcorr[S, R, N] <- res$cov.boot.wcorr  
      cov.pbend[S, R, N] <- res$cov.boot.pbcor
      cov.spear[S, R, N] <- res$cov.boot.spear 
      cov.pears[S, R, N] <- res$cov.boot.pears
      cov.pears.zou[S, R, N] <- res$cov.zou.pears
    }
  }
save(res.spear, res.pears.boot, res.pears.fish, 
     res.pears.hc4, res.wcorr, res.pbend,
     cov.wcorr, cov.pbend, cov.spear, cov.pears, cov.pears.zou
     rhoseq, nseq, nsim, es, g, h,
     file = "./data/diff01_g1h02_varyrho.RData")
}
beep(8)
```
