---
title: "Compare independent correlations: rho = 0.3, 0 difference, vary g"
author: "Guillaume A. Rousselet"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: no
    number_sections: no
    toc: yes
    toc_depth: 2
---

# Dependencies
```{r}
library(ggplot2)
library(tibble)
library(beepr)
source("./functions/theme_gar.txt")
## HC4 method to compare two independent Pearson's correlations
source("./functions/hc4.txt")
## g & h multivariate distributions
source("./functions/ghmult.txt")
## bootstrap + core corr functions
source("./functions/corfun.txt")
## meta-function for simulations
source("./functions/twocor.all.txt")
## plot functions
source("./functions/make_figures.txt")
```

```{r}
sessionInfo()
```

# h = 0
```{r, eval = FALSE}
gseq <- seq(0, 1, 0.1)
ng <- length(gseq)
h <- 0
nseq <- seq(50, 500, 50)
nmax <- max(nseq)
nsim <- 5000

rho <- 0.3
es <- 0 # effect size
cmat1 <- matrix(c(1,rho,rho,1),2,2)
cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)

res.wcorr <- array(0, dim = c(nsim, ng, length(nseq)))
res.pbend <- array(0, dim = c(nsim, ng, length(nseq)))
res.spear <- array(0, dim = c(nsim, ng, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, ng, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, ng, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, ng, length(nseq)))

for(G in 1:ng){
  beep(2)
  set.seed(21) # set seed inside G loop to allow direct comparisons
  # (numbers are the same except for the skewness)
  g <- gseq[G]
  print(paste("sim rho = 0.3, diff = 0, h = 0, g =",g))
  for(S in 1:nsim){
    if(S %% 1000 == 0){print(paste0("sim ",S," / ",nsim,"..."))}
    # generate max n sample
    out1 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat1)  
    out2 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat2) 
  for(N in 1:length(nseq)){
    # subsample from max n sample
      res <- twocor.all(out1[1:nseq[N],1],out1[1:nseq[N],2],out2[1:nseq[N],1],out2[1:nseq[N],2])
      res.wcorr[S, G, N] <- res$sig.boot.wcorr
      res.pbend[S, G, N] <- res$sig.boot.pbcor
      res.spear[S, G, N] <- res$sig.boot.spear
      res.pears.boot[S, G, N] <- res$sig.boot.pears
      res.pears.fish[S, G, N] <- res$sig.fisher.pears
      res.pears.hc4[S, G, N] <- res$sig.hc4.pears
    }
  }
save(res.spear, res.pears.boot, res.pears.fish, 
     res.pears.hc4, res.wcorr, res.pbend,
     gseq, nseq, nsim, es, h, rho,
     file = "./data/rho03_diff00_h0_varyg.RData")
}
beep(8)
```

## Results

### Load data + generate all plot objects
```{r}
load("./data/rho03_diff00_h0_varyg.RData")
PO <- plot.all.fp.g(ylim = c(0, 0.2))
```

### Pearson + Fisher
```{r}
pA <- PO$p.pears.fish 
pA
```

### Pearson + bootstrap
```{r}
pB <- PO$p.pears.boot + guides(col = guide_legend(ncol = 3, override.aes = list(size = 3)))
pB 
```

### Pearson + HC4
```{r}
pC <- PO$p.pears.hc4
pC
```

### Spearman + bootstrap
```{r}
pD <- PO$p.spear
pD
```

### Percentage bend + bootstrap
```{r}
pE <- PO$p.pbend
pE
```

### Winsorized correlation + bootstrap
```{r}
pF <- PO$p.wcorr
pF
```

### Combine panels

```{r, eval = FALSE}
cowplot::plot_grid(pA, pB, pC, pD, pE, pF,
                   labels = c("A", "B", "C", "D", "E", "F"))
ggsave("./figures/rho03_diff00_h0_vg.pdf", width = 20, height = 30, units = "cm")
```

# h = 0.2
```{r, eval = FALSE}
gseq <- seq(0, 1, 0.1)
ng <- length(gseq)
h <- 0.2
nseq <- seq(50, 500, 50)
nmax <- max(nseq)
nsim <- 5000
rho <- 0.3
es <- 0 # effect size
cmat1 <- matrix(c(1,rho,rho,1),2,2)
cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)

res.wcorr <- array(0, dim = c(nsim, ng, length(nseq)))
res.pbend <- array(0, dim = c(nsim, ng, length(nseq)))
res.spear <- array(0, dim = c(nsim, ng, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, ng, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, ng, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, ng, length(nseq)))

for(G in 1:ng){
  beep(2)
  set.seed(21) # set seed inside R loop to allow direct comparisons
  # (numbers are the same except for the correlation)
  g <- gseq[G]
  print(paste("sim rho = 0.3, diff 00, h=0, g =",g))
  for(S in 1:nsim){
    if(S %% 1000 == 0){print(paste0("sim ",S," / ",nsim,"..."))}
    # generate max n sample
    out1 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat1)  
    out2 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat2) 
  for(N in 1:length(nseq)){
    # subsample from max n sample
      res <- twocor.all(out1[1:nseq[N],1],out1[1:nseq[N],2],out2[1:nseq[N],1],out2[1:nseq[N],2])
      res.wcorr[S, G, N] <- res$sig.boot.wcorr
      res.pbend[S, G, N] <- res$sig.boot.pbcor
      res.spear[S, G, N] <- res$sig.boot.spear
      res.pears.boot[S, G, N] <- res$sig.boot.pears
      res.pears.fish[S, G, N] <- res$sig.fisher.pears
      res.pears.hc4[S, G, N] <- res$sig.hc4.pears
    }
  }
save(res.spear, res.pears.boot, res.pears.fish, 
     res.pears.hc4, res.wcorr, res.pbend,
     gseq, nseq, nsim, es, h, rho,
     file = "./data/rho03_diff00_h02_varyg.RData")
}
beep(8)
```

## Results

### Load data + generate all plot objects
```{r}
load("./data/rho03_diff00_h02_varyg.RData")
PO <- plot.all.fp(ylim = c(0, 0.2))
```

### Pearson + Fisher
```{r}
pA <- PO$p.pears.fish + guides(col = guide_legend(ncol = 3, override.aes = list(size = 3)))
pA
```

### Pearson + bootstrap
```{r}
pB <- PO$p.pears.boot
pB
```

### Pearson + HC4
```{r}
pC <- PO$p.pears.hc4
pC
```

### Spearman + bootstrap
```{r}
pD <- PO$p.spear
pD
```

### Percentage bend + bootstrap
```{r}
pE <- PO$p.pbend
pE
```

### Winsorized correlation + bootstrap
```{r}
pF <- PO$p.wcorr
pF
```

### Combine panels

```{r, eval = FALSE}
cowplot::plot_grid(pA, pB, pC, pD, pE, pF,
                   labels = c("A", "B", "C", "D", "E", "F"))
ggsave("./figures/rho03_diff00_h02_vg.pdf", width = 20, height = 30, units = "cm")
```

